---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# AKaerial

<!-- badges: start -->
<!-- badges: end -->

The goal of AKaerial is to provide a transparent and repeatable workflow to generate, visualize, and report out estimates from the suite of aerial surveys undertaken by the Division of Migratory Bird Management (MBM) in the Alaska Region of the US Fish and Wildlife Service.  Pilots and observers in the Alaska Region have been collecting aerial survey data on waterfowl for over 50 years.  In general, a pilot and observer fly low-level transects over a designated habitat-based study area (usually breeding grounds) and record observations of single drakes (or single birds for monomorphic species), breeding pairs, and flocks. Observations are summarized as densities, and densities are extrapolated to population indices.  Resultant indices are sometimes used in harvest management decisions by both state and federal agencies.  AKaerial was created not only to streamline these steps, but to provide a transparent, repeatable, and defensible alternative to past estimate-generating processes.  

The most likely use of AKaerial would be to take advantage of the historic tables of estimates for projects on the Arctic Coastal Plain ( [ACPHistoric](../help/ACPHistoric), 2007-2019), Copper River Delta ( [CRDHistoric](../help/CRDHistoric), 1986-2019), and Yukon-Kuskokwim River Delta for ducks ( [YKDHistoric](../help/YKDHistoric), 1988-2019) and geese ( [YKGHistoric](../help/YKGHistoric), 1985-2019).  AKaerial includes figure-, table-, and map-generating functions that are designed to visualize data, design files, and estimates as a quality control step as well as for the production of reports, presentations, and manuscripts.

AKaerial also includes a function ( [GreenLight](../help/GreenLight)) that provides a rigorous and documented quality control process.  This results in a structured and consistent treatment of both raw transcribed and archived data files throughout the data life cycle.   

## Bugs, Issues, Feature Requests

AKaerial is under constant development. It will have occasional bugs and other issues.  Please report any issues or feature requests on the [AKaerial Github repository](https://github.com/USFWS/AKaerial/issues). 

## Installation

You can install the released version of AKaerial from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("USFWS/AKaerial", ref = "master", build_vignettes = TRUE)
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("USFWS/AKaerial", ref = "development", build_vignettes = TRUE)
```
Then explore the vignettes with:

```r
vignette("IntroToAKaerial", package="AKaerial")
```

## Examples

Below are several examples of functions in AKaerial that check, process, and summarize data and observations.


To visualize historic estimates as tables or figures:

```{r example, echo = TRUE, message=FALSE, warning = FALSE}
library(AKaerial)

ReportFigure(data=YKGHistoric$combined, 
             species="EMGO", 
             index=c("total", "total.var", "itotal", "itotal.var", "ibb", "ibb.var"),
             year=c(2007:2019), 
             title="Emperor goose population indices from the \nYukon-Kuskokwim Delta, 2007-2019", 
             x.label = "Year", 
             y.label="Index Estimate", 
             leg.values = c(3,2,1), 
             leg.labels = c("Total", "Indicated Total", "Indicated Breeding"), 
             leg.limits = c("total", "itotal", "ibb"),
             test.out = FALSE)




ReportFigure(data=ACPHistoric$combined, 
             species="SPEI", 
             index=c("itotal", "itotal.var"),
             year=c(2007:2019),
             yr.avg=3,
             title="Indicated total spectacled eiders (and 3-year average) on the \nArctic Coastal Plain, 2007-2019", 
             x.label = "Year", 
             y.label="Indicated Total", 
             leg.values = c(2,1), 
             leg.labels = c("Indicated Total", "3-year Avg"), 
             leg.limits = c("itotal", "avg1"),
             test.out = FALSE)


ReportTable(data=YKDHistoric$combined, 
            species="SPEI", 
            index=c("itotal", "itotal.var"),
            year=c(1988:2019), 
            yr.avg=3, 
            cap="Indicated total spectacled eiders on the Yukon-Kuskokwim Delta, 1988-2019.", 
            new.names=c("Year", "Indicated Total", "SE", "3-yr Avg", "SE"))


ReportTable(data=CRDHistoric$combined, 
            species="DCGO", 
            index=c("total", "total.var", "itotal", "itotal.var"),
            year=c(1995:2019), 
            yr.avg=3, 
            cap="Total and indicated total dusky Canada geese on the Copper River Delta, 1995-2019.", 
            new.names=c("Year", "Total", "SE", "3-yr Avg", "SE", "Indicated Total", "SE", "3-yr Avg", "SE"))

```

