---
title: "Aerial Survey Summary Report"
author: "Chuck Frost"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)

library(leaflet)
library(plyr)
library(dplyr)
library(kableExtra)
library(DT)

```

## Introduction

This document provides a combined index estimate for 2 observers on Region 7 MBM Waterfowl Division aerial surveys.  Currently supported surveys include the Arctic Coastal Plain (ACP), Yukon-Kuskokwim Delta duck (YKD) and goose (YKG), Copper River Dusky (CRD), Black Scoter Survey (BLSC), and North American Waterfowl Breeding Population (BPOP).

All data checks are currently using assumptions for data fields as expected to be written into protocols and summarized and analyzed using the R7MBM survey analysis R package **AKaerial**.  For information and current version, see <https://github.com/cfrost3/AKaerial>.

You entered the function call:

CombinedReport(strata.path = `r strata.path`, 
  transect.path = `r transect.path`, 
  data1.path = `r data1.path`,
  data2.path = `r data2.path`,
  area = `r area`,
  threshold = `r threshold`)

You have chosen the files:
```{r files}

print(data1.path)

print(data2.path)

```

For the `r area` survey in `r data1$obs$Year[1]`. There were `r length(strata.names)` strata detected and maps appear below with transect design and observations added.  


```{r maps, echo=FALSE}
createMaps <- function(strata.name){
mymap <- leaflet() %>%
addTiles() %>%
  addPolygons(data=split(strata.proj, strata.proj$STRATNAME)[[strata.name]],
                fillOpacity=.4,
                stroke=TRUE,
                color="white",
                opacity=1,
                weight=1,
                popup = paste("Strata: ", strata.proj$STRATNAME)) %>%
  
  addPolylines(data=split(split.design, split.design$STRATNAME)[[strata.name]],
                 color="black",
                 weight=4,
                 opacity=.9,
                 label=~split.design$OBJECTID,
                 popup = paste("Strata: ", split.design$STRATNAME, "<br>",
                               "Old Transect: ", split.design$ORIGID, "<br>",
                               "New Transect: ", split.design$OBJECTID, "<br>",
                               "Split Transect: ", split.design$SPLIT, "<br>",
                               "Length: ", split.design$len))  %>%
  
  addCircleMarkers(data=total.obs[as.character(total.obs@data$ctran) %in% as.character(split(split.design, split.design$STRATNAME)[[strata.name]]$SPLIT),],
                     radius = 3,
                     color = "red",
                     stroke = FALSE,
                     fillOpacity = 1
                  
                  ) %>%

    leaflet::addProviderTiles("Esri.WorldImagery")
  
return(list(tags$h1(strata.name), mymap))
}

htmltools::tagList(lapply(as.list(names(split(strata.proj, strata.proj$STRATNAME))), function(x) createMaps(x) ))


   
```








## Diagnostics

The following table shows the stratification names and areas in square kilometers as derived from the file `r strata.path`. 

```{r tables1, echo=FALSE}

datatable(data1$strata[,1:2])


```









The following table displays the transects (Original is the design file or navigational name, PartOf is the corrected name based on latitude/longitude, extraneous pieces, misnumbering, etc.), their sampled areas (SampledArea, in square kilometers), Length in kilometers, and associated Strata.  


```{r tables2, echo=FALSE}

datatable(rbind(data1$flight, data2$flight))

```








The following table provides a summary of the sampled area by observer and strata:

```{r tables3, echo=FALSE}

datatable(aggregate(SampledArea~Observer+Strata, data=rbind(data1$flight, data2$flight), FUN=sum))

```









## Results

If the above diagnostic checks are reasonable, the following observer-specific index estimates apply:

```{r results1, echo=FALSE}

datatable(rbind(obs1.est$estimates, obs2.est$estimates))

```





For more detailed results by observer:

```{r results2, echo=FALSE}

datatable(rbind(obs1.est$counts.final, obs2.est$counts.final))

```





For combined estimates:

```{r results3, echo=FALSE}

datatable(combine.est)

```





And finally, combined stratified estimates:

```{r results4, echo=FALSE}

datatable(est.by.strata)

```






