---
title: "AKaerial Data Quality"
author: "Chuck Frost"
date: "July 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This document provides a preliminary data quality check for data collected by Region 7 MBM Waterfowl Division aerial surveys.  Currently supported surveys include the Arctic Coastal Plain (ACP), Yukon-Kuskokwim Delta duck (YKD) and goose (YKG), Copper River Dusky (CRD), and North American Waterfowl Breeding Population (BPOP).

All data checks are currently using assumptions for data fields as expected to be written into protocols and summarized and analyzed using the R7MBM survey analysis R package **AKaerial**.  For information and current version, see <https://github.com/cfrost3/AKaerial>.


You have chosen the file:
```{r file}

path.name

```

For the area:

```{r area}

area

```


## Spatial

There were `r n.latlong` spatial coordinate warnings.  These are longitude or latitude values that fall outside the normal boundaries of the survey. 

`r if(s.latlong != "green") {"\\begin{comment}"}`
Status - GREEN
`r if(s.latlong != "green") {"\\end{comment}"}`

`r if(s.latlong != "yellow") {"\\begin{comment}"}`
Status - YELLOW
`r if(s.latlong != "yellow") {"\\end{comment}"}`



