---
title: "AKaerial Data Quality"
author: "Chuck Frost"
date: "December, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This document provides a preliminary data quality check for data collected by Region 7 MBM Waterfowl Division aerial surveys.  Currently supported surveys include the Arctic Coastal Plain (ACP), Yukon-Kuskokwim Delta duck (YKD) and goose (YKG), Copper River Dusky (CRD), and North American Waterfowl Breeding Population (BPOP).

All data checks are currently using assumptions for data fields as expected to be written into protocols and summarized and analyzed using the R7MBM survey analysis R package **AKaerial**.  For information and current version, see <https://github.com/cfrost3/AKaerial>.


You have chosen the file:
```{r file}

print(path.name)

```

For the area:

```{r area}

print(area)

```

We have standardized the mandatory column headings as:

`r necessary`

`r if(s.colmatch != "green") {"\\begin{comment}"}`
This file contains all of the required column names.

Status - GREEN
`r if(s.colmatch != "green") {"\\end{comment}"}`

`r if(s.colmatch != "red") {"\\begin{comment}"}`
This file does not contain all of the required column names.

`r colnames(data)`

Status - RED
`r if(s.colmatch != "red") {"\\end{comment}"}`

## Spatial

There were `r n.latlong` spatial coordinate warnings.  These are longitude or latitude values that fall outside the normal boundaries of the survey. If these values are attributed to a transect used in the ratio estimate and cannot be interpolated, they may be removed.

`r if(s.latlong != "green") {"\\begin{comment}"}`
Status - GREEN
`r if(s.latlong != "green") {"\\end{comment}"}`

`r if(s.latlong != "yellow") {"\\begin{comment}"}`
Status - YELLOW

This indicates a potential error in `r n.latlong` of the latitude/longitude values, specified below:

`r latlong`

`r if(s.latlong != "yellow") {"\\end{comment}"}`

## Swans

Swan nests should only be recorded if a swan is also present (active).  Correct notation is recording the nest as open, 1 and the associated swan on a separate entry as single or pair, 1.

`r if(s.swan != "green") {"\\begin{comment}"}`
All swan nests and associated swans appear to be recorded correctly.

Status - GREEN
`r if(s.swan != "green") {"\\end{comment}"}`

`r if(length(test.swan$bad[,1]) == 0) {"\\begin{comment}"}`
The following nests are not recorded as open, 1:

`r print(test.swan$bad)`

Status - RED
`r if(length(test.swan$bad[,1]) == 0) {"\\end{comment}"}`

`r if(length(test.swan$missing[,1]) == 0) {"\\begin{comment}"}`
The following nests do not have associated swan records:

`r print(test.swan$missing)`

Status - RED
`r if(length(test.swan$missing[,1]) == 0) {"\\end{comment}"}`






