---
title: "AKaerial Analysis File Summary"
author: "Chuck Frost"
email: "charles_frost@fws.gov"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plyr)
```

## Introduction

This document provides a log file of changes made to raw observer data collected by Region 7 MBM Waterfowl Division aerial surveys in the creation of a file suitable for standard ratio estimation in the package **AKaerial**.  Historical data must meet quality checks against the current data dictionary for R7 MBM aerial surveys (found at <https://github.com/cfrost3/AKaerial/tree/master/data/AKAerialDataDictionary.csv>).

Currently supported surveys include the Arctic Coastal Plain (ACP), Yukon-Kuskokwim Delta duck (YKD) and goose (YKG), Copper River Dusky (CRD), and North American Waterfowl Breeding Population (BPOP).

All data checks are currently using assumptions for data fields as expected to be written into protocols and summarized and analyzed using the R7MBM survey analysis R package **AKaerial**.  For information and current version, see <https://github.com/cfrost3/AKaerial>.


You have chosen the file:
```{r file}

print(path.name)

```

For the area:

```{r area}

print(area)

```

`r if(!("Swan" %in% fix)) {"\\begin{comment}"}`
## Swans

Swan nests should only be recorded if a swan is also present (active).  Correct notation is recording the nest as open, 1 and the associated swan on a separate entry as single or pair, 1.

There were `r length(test.swan$bad[,1])` swan nests not recorded as open, 1 that were changed to open, 1.

There were `r length(test.swan$missing[,1])` nests without associated swan records.  Identical records were created for each nest as either pair, 1 (when the nest was entered as pair, 1) or single, 1 (when the nest was entered as single, 1).


`r if(!("Swan" %in% fix)) {"\\end{comment}"}`


`r if(!("Obs_Type" %in% fix)) {"\\begin{comment}"}`

## Obs_Type (Unit) Codes

The following Obs_Type codes are acceptable- single, pair, open, flkdrake.  Leading or trailing spaces are not acceptable.  From the inception of most surveys through 2014, we noticed a tendency of some observers to incorrectly record all Obs_Type codes as either open or NA.  In these cases, a correction is applied to the raw Obs_Type codes to reflect the correct treatment of Obs_Type in the data.  This file was flagged as needing an Obs_Type correction.  The correction is as follows:

If Num = 1; Obs_Type = single
If Num = 2; Obs_Type = pair, Num = 1
If Num >= 3; Obs_Type = open
If the record is a swan nest; Obs_Type = open, Num = 1

Any flocked drakes were recorded on multiple lines sharing the same Lon, Lat as open, 1. For monomorphic species, open with Num >= 3 indicates a flock. 

`r if(!("Obs_Type" %in% fix)) {"\\end{comment}"}`


`r if(!("Seat" %in% fix)) {"\\begin{comment}"}`

## Seat Codes

Seat codes must be either LF, RF, LR, or RR, and are case-sensitive.  The seat codes in the raw file were input incorrectly and were corrected for this version.

`r if(!("Seat" %in% fix)) {"\\end{comment}"}`

`r if(!("Observer" %in% fix)) {"\\begin{comment}"}`

## Observers

The observer initials must be uppercase.  Observer initials in the raw file were input incorrectly and were corrected for this version.

`r if(!("Observer" %in% fix)) {"\\end{comment}"}`

`r if(!("Species" %in% fix)) {"\\begin{comment}"}`

## Species Codes
The current list of acceptable species codes can be found at:
<https://github.com/cfrost3/AKaerial/tree/master/data/AKAerialSpeciesAOUCodes.csv>

The following species codes are not acceptable and were updated:

```{r spec_missing}
kableExtra::kable_styling(kableExtra::kable(table(test.species$change$Species)))
```

Any species code marked as XXXX in the QAQC column of the species file is considered to be outside the design of the survey and not pertinent to the project.  These observations were removed from the raw file before this version was finalized.  Any observations in this category in the raw data should be considered anecdotal and not collected in any scientific manner.  These observations were not collected consistently by observer, year, survey, day, or species.  

`r if(!("Species" %in% fix)) {"\\end{comment}"}`

